Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-05-18 paper.tex
edits motivated by Dan's comments

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing primitives
  that explain an image. This set of drawing primitives is like an execution trace for a graphics program. From this trace we use
  program synthesis techniques to recover a graphics
  program with constructs like variable bindings, iterative loops, or
  simple kinds of conditionals. With a graphics program in hand,
  we can correct errors made by the deep network, extrapolate drawings, and cluster drawings by use of similar high-level geometric structures.
  Taken together these results are a step towards
  agents that induce useful, human-readable programs from perceptual input.  
\end{abstract}

@@

The CogSketch~\cite{forbus2011cogsketch} system also aims to have a high-level understanding of hand-drawn figures. Their primary goal is cognitive modeling, whereas we are interested in building an automated AI application.

%% open domain sketch understanding. structure mapping style model. also taps into cyc? the user segments the ink and labels them with concepts from knowledge base. like us they focus on compositional understanding. ``cognitive simulation", solves things like Raven matrices in IQ tests. not really trying right now to be automated like our system.

@@

\remark{Do you also want to cite your own work on ``Unsupverised Learning by Program Synthesis'' here?}

@@

  \node[draw,ultra thick,minimum size = 1cm](c1) at ([xshift = 1.5cm]features.north) {MLP};
  \node(l1) at ([yshift = -2cm]c1.south) {\verb|circle(|};
  \node[draw,ultra thick,minimum size = 1cm](c2) at ([xshift = 1cm]c1.east) {MLP};
  \draw[->,ultra thick,red] (c1.south) -- (l1.north);
  \node(l2) at ([yshift = -2cm]c2.south) {\verb|X=3,|};
  \draw[->,ultra thick,red] (c2.south) -- (l2.north);
  \node[draw,ultra thick,minimum size = 1cm](c3) at ([xshift = 1cm]c2.east) {MLP};
  \node(l3) at ([yshift = -2cm]c3.south) {\verb|Y=14)|};
  \draw[->,ultra thick,red] (c3.south) -- (l3.north);

@@

\begin{figure}[H]
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-10.png}
    \includegraphics[width = 2cm]{figures/10-parse.png}    
    \subcaption{Noisy input}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-21.png}
    \includegraphics[width = 2cm]{figures/21-parse.png}    
    \subcaption{A graphical model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-77.png}
    \includegraphics[width = 2cm]{figures/77-parse.png}    
    \subcaption{A figure from a deep learning textbook}
    \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-0.png}
    \includegraphics[width = 2cm]{figures/0-parse.png}    
    \subcaption{Near miss}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-38.png}
    \includegraphics[width = 2cm]{figures/38-parse.png}    
    \subcaption{Failing on a Ising model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-34.png}
    \includegraphics[width = 2cm]{figures/34-parse.png}     
    \subcaption{Illusory contours}
    \end{minipage}
  \caption{Example drawings above model outputs. See also Fig.~\ref{firstPageExamples}}\label{lotsOfHandDrawings}~\remark{Showing `failure cases' (the last three above) out-of-context seems not great. Perhaps remind people in the caption that stochastic search can help correct some of these issues?}
\end{figure}
For each drawing we annotated a ground truth trace, and evaluated the
model by asking it to sample many candidate traces for each drawing.
For 57\% of the drawings the Top-1 most likely sample exactly matches the
ground truth; as we consider more samples the model encounters traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
Because our current model sometimes makes mistakes on hand drawings,
we envision the current system working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

@@

We constrain the space of allowed programs by writing down a context
free grammar over a space of programs. Although it might be desirable
to synthesize programs in a Turing-complete language such as Lisp or
Python, a more tractable approach is to specify what in the program
languages community is called a Domain Specific Language (DSL)~\remark{citation?}. Our DSL (Table~\ref{DSL})
encodes prior knowledge of what graphics programs tend to look like.

@@

The constrained optimization problem in
equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to
program synthesis problems like these. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see supplement.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program 's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem, and then uses a SAT solver to find a
minimum cost program satisfying the specification.  In exchange for
not having any guarantees on how long it will take to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.

Why synthesize a graphics program,
if the execution trace already suffices to recover the objects in an image?
Within our domain of hand-drawn figures, graphics program synthesis has several uses:
\remark{I'm of two minds about how these subsections should be ordered. The current ordering leads with the coolest results, which is nice. But it's a bit...deflating?...to start with such cool results and then end on the somewhat technical point of how the synthesizer can help correct parse errors. An alternative would be to flip the ordering, starting with the technical point and building toward progressively cooler results. This requires a bit more patience on the part of the reader, but the overall narrative flow/payoff feels better. In any case, whatever ordering you pick should be the same order as these applications are mentioned in the abstract.}

@@

\subsection{Correcting errors made by the parsing network}
The program synthesizer can help correct errors from the parse proposal network by favoring execution traces which lead to more concise or general programs.
For example, one generally prefers figures with perfectly aligned objects over figures whose parts are slightly misaligned -- and precise alignment lends itself to short programs.
Similarly, figures often have repeated parts,
which the program synthesizer might be able to model as a loop or reflectional symmetry.
So, in considering several candidate traces proposed by the neural network,
we might prefer traces whose best programs have desirable features such being short or having iterated structures.

>>>>

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing primitives
  that explain an image. This set of drawing primitives is like an execution trace for a graphics program. From this trace we use
  program synthesis techniques to recover a graphics
  program with constructs like variable bindings, iterative loops, or
  simple kinds of conditionals. With a graphics program in hand,
  we can correct errors made by the deep network, cluster drawings by use of similar high-level geometric structures, and extrapolate drawings.
  Taken together these results are a step towards
  agents that induce useful, human-readable programs from perceptual input.  
\end{abstract}

@@

@@

The CogSketch~\cite{forbus2011cogsketch} system also aims to have a
high-level understanding of hand-drawn figures. Their primary goal is
cognitive modeling (eg, they apply their system to solving IQ-test
style visual reasoning problems), whereas we are interested in
building an automated AI application (eg, in our system the user need
not annotate which strokes correspond to which shapes; our neural
network produces something equivalent to the annotations).  A key
similarity however is that both CogSketch and our system have as a
goal to make it easier to produce nice-looking figures.
Unsupervised Program Synthesis~\cite{ellis2015unsupervised} is a related framework which was also applied to geometric reasoning problems. The goals of~\cite{ellis2015unsupervised} were cognitive modeling,
and they applied their technique to synthetic scenes used in human behavioral studies.
%% open domain sketch understanding. structure mapping style model. also taps into cyc? the user segments the ink and labels them with concepts from knowledge base. like us they focus on compositional understanding. ``cognitive simulation", solves things like Raven matrices in IQ tests. not really trying right now to be automated like our system.

@@

  \node[draw,ultra thick,minimum size = 1cm](c1) at ([xshift = 1.5cm]features.north) {MLP};
  \node(l1) at ([yshift = -2cm]c1.south) {\verb|circle(|};
  \node[draw,ultra thick,minimum size = 1cm](c2) at ([xshift = 1cm]c1.east) {MLP};
  \draw[->,ultra thick,red] (c1.south) -- (l1.north);
  \node(l2) at ([yshift = -2cm]c2.south) {\verb|X=7,|};
  \draw[->,ultra thick,red] (c2.south) -- (l2.north);
  \node[draw,ultra thick,minimum size = 1cm](c3) at ([xshift = 1cm]c2.east) {MLP};
  \node(l3) at ([yshift = -2cm]c3.south) {\verb|Y=12)|};
  \draw[->,ultra thick,red] (c3.south) -- (l3.north);

@@

\begin{figure}[H]
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-10.png}
    \includegraphics[width = 2cm]{figures/10-parse.png}    
    \subcaption{Noisy input}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-21.png}
    \includegraphics[width = 2cm]{figures/21-parse.png}    
    \subcaption{A graphical model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-77.png}
    \includegraphics[width = 2cm]{figures/77-parse.png}    
    \subcaption{A figure from a deep learning textbook}
    \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-0.png}
    \includegraphics[width = 2cm]{figures/0-parse.png}    
    \subcaption{Near miss}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-38.png}
    \includegraphics[width = 2cm]{figures/38-parse.png}    
    \subcaption{Failing on a Ising model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-34.png}
    \includegraphics[width = 2cm]{figures/34-parse.png}     
    \subcaption{Illusory contours}
    \end{minipage}
  \caption{Example drawings above model outputs. See also Fig.~\ref{firstPageExamples}. Stochastic search (SMC) can help correct for these errors, as can the program synthesizer (Section~\ref{synthesizerHelpsParsing})}\label{lotsOfHandDrawings}%~\remark{Showing `failure cases' (the last three above) out-of-context seems not great. Perhaps remind people in the caption that stochastic search can help correct some of these issues?}
\end{figure}
For each drawing we annotated a ground truth trace, and evaluated the
model by asking it to sample many candidate traces for each drawing.
For 57\% of the drawings the Top-1 most likely sample exactly matches the
ground truth; as we consider more samples the model encounters traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
Because our current model sometimes makes mistakes on hand drawings,
we envision the current system working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

@@

We constrain the space of allowed programs by writing down a context
free grammar over a space of programs. Although it might be desirable
to synthesize programs in a Turing-complete language such as Lisp or
Python, a more tractable approach is to specify what in the program
languages community is called a Domain Specific Language (DSL)~\cite{polozov2015flashmeta}. Our DSL (Table~\ref{DSL})
encodes prior knowledge of what graphics programs tend to look like.

@@

The constrained optimization problem in
equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to
program synthesis problems like these. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see supplement.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem, and then uses a SAT solver to find a
minimum cost program satisfying the specification.  In exchange for
not having any guarantees on how long it will take to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.

Why synthesize a graphics program,
if the execution trace already suffices to recover the objects in an image?
Within our domain of hand-drawn figures, graphics program synthesis has several uses:
%\remark{I'm of two minds about how these subsections should be ordered. The current ordering leads with the coolest results, which is nice. But it's a bit...deflating?...to start with such cool results and then end on the somewhat technical point of how the synthesizer can help correct parse errors. An alternative would be to flip the ordering, starting with the technical point and building toward progressively cooler results. This requires a bit more patience on the part of the reader, but the overall narrative flow/payoff feels better. In any case, whatever ordering you pick should be the same order as these applications are mentioned in the abstract.}

\subsection{Correcting errors made by the neural network}\label{synthesizerHelpsParsing}
The program synthesizer can help correct errors from the execution trace proposal network by favoring execution traces which lead to more concise or general programs.
For example, one generally prefers figures with perfectly aligned objects over figures whose parts are slightly misaligned -- and precise alignment lends itself to short programs.
Similarly, figures often have repeated parts,
which the program synthesizer might be able to model as a loop or reflectional symmetry.
So, in considering several candidate traces proposed by the neural network,
we might prefer traces whose best programs have desirable features such being short or having iterated structures.

@@
